---
title: "Tests of Independence"
author: "Le Wang"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This program illustrates how to perform tests of independence. 

```{r}
library(MASS)
head(survey)
```

## Example 1: Smoking habit and Exercise Levels

```{r}
mytable <- table(survey$Smoke, survey$Exer)
mytable
```


Lets perform the $\chi^2$ test of independence:
```{r}
chisq.test(mytable)
```



The warning message found in the solution above is due to the small cell values in the contingency table. We can combine those small cells, for example,

```{r}
mytable<-cbind(mytable[,"Freq"],mytable[,"None"]+ mytable[,"Some"])
mytable
```

```{r}
chisq.test(mytable)
```



## Example 2: Super Bowl and White Jerseys (in year 2018)


Lets manually input the data in superbowl outcomes
```{r}
mytable <- matrix(c(33,19,19,33),ncol=2,byrow=TRUE)
mytable<-as.table(mytable)
mytable
```


```{r}
chisq.test(mytable)
```


# Alternative Tests: Fishers Exact Test

```{r}
fisher.test(mytable)
```



# Cochran-Mantel-Haenszel Test of conditional independence (will be discussed later)
# mantelhaen.test(mytable)