---
title: "Conditional Distribution for More than Two Variables"
output: html_notebook
---

Our goal is eventually use the data with information `race`, `gender`, and `age` to calculate the following conditional distribution

$$
\Pr[\text{Race and Gender} \mid \text{Age}] = \frac{\Pr[\text{Race, Gender, and Age}]}{\Pr[\text{Age}]}
$$

Let's read our data into `R` first. You need to create a folder called `data` and save the files into this folder. Make sure your program is in the same folder as the data folder. 

```{r}
cnames<-read.csv("data/cnames.csv") 
FLvoters<-read.csv("data/FLvoters.csv")
```


Let's omit the missing values in the dataset 

```{r}
FLvoters <- na.omit(FLvoters)
```

Let's examine the dataset
```{r}
str(FLvoters)
```



As we can see, `age` is a continuous variable. Let's discretize it first so that we can directly apply our formula. 

```{r}
FLvoters$age.group<- NA
FLvoters$age.group[FLvoters$age <= 20] <- 1
FLvoters$age.group[FLvoters$age > 20 & FLvoters$age <= 40] <- 2
FLvoters$age.group[FLvoters$age > 40 & FLvoters$age <= 60] <- 3
FLvoters$age.group[FLvoters$age >60] <- 4
```

Or, you can use the tidy approach:

```{r}
library(tidyverse)
FLvoters <- FLvoters %>% 
  mutate(
    age.group2 = case_when(
      age <= 20 ~ 1,
      age >20 & age<=40  ~ 2,
      age >40 & age<=60  ~ 3,
      age>60 ~ 4
    )
  )
```

Lets check if these two approaches are equilvalent. 

```{r}
table(FLvoters$age.group2, FLvoters$age.group)
```



```{r}
summary(FLvoters)
```


Now all the variables of interest are discrete. The process is exactly the same. Just now, we need to put three variables in our `table()` command and compute the joint distribution of all three. 

```{r}
mytable<-table(race=FLvoters$race,
               age.group=FLvoters$age.group, 
               gender=FLvoters$gender)
```

```{r}
joint.dist <- prop.table(mytable)
```

The joint distribution is inevitably more complicated

```{r}
joint.dist
```

Remember that in order to calculate the conditional distribution, we also need the marginal distribution of the conditioning variable(s). 

But the following code is


Not as convenient


!!!!

```{r}
addmargins(joint.dist)
```

Why? Because it calculates the marginal distribution for each sub-table as well. 

```{r}
margin.dist.age <- prop.table(table(FLvoters$age.group))
margin.dist.age
```

Suppose that we are interested in 

$$
\begin{aligned}
\Pr[\text{black and female}|\text{above 60}] &=\frac{\Pr[\text{Black and above 60 and female}]}{\Pr[\text{above 60}]}\\
 &=\frac{0.0188741358}{0.35092725}
\end{aligned}
$$

```{r}
joint.dist["black",4,"f"]/margin.dist.age[4]
```

Or, we can calculate the following conditional distribution for all race and gender combinations by not specifying the first and third dimensions!

$$
\begin{aligned}
\Pr[\text{Race and Gender}|\text{above 60}] &=\frac{\Pr[\text{Race and above 60 and Gender}]}{\Pr[\text{above 60}]}
\end{aligned}
$$

```{r}
joint.dist[,4,]/margin.dist.age[4]
```

