---
title: "An Illustration of Bayesian Estimation"
output: html_notebook
---

This program illustrates an exmaple of Bayesian Estimation using R. This is adapted from Kruschke's book **Doing Bayesian Data Analysis**. 

Recall the Bayesian Rule:

$$
f(\theta|x)=\frac{f(x|\theta)\cdot f(\theta)}{f(x)}
$$

Step 1. Let's define our prior

$$
f(\theta)
$$


```{r}
  Theta <- c(.25, .5, .75)
  Theta
  
# pTheta is the vector of prior probabilities on the theta values.

  pTheta <- c(.25, .5, .25) # Makes a triangular belief distribution.
  sum(pTheta) # Makes sure that beliefs sum to 1.  
```

Step 2. Let's look at our data (evidence)

```{r}
  Data <- c(rep(1,3),rep(0,9))
  Data
  nHeads <- sum( Data )
  nTails <- length( Data ) - nHeads
  
  nHeads
  nTails
```

What is the likelihood of evidence for each $\theta$

$$
f(x\mid \theta)
$$


```{r}
  pDataGivenTheta = Theta^nHeads * (1-Theta)^nTails

  pDataGivenTheta
```

What is the probability of our data?
$$
f(x)  =f(x|\theta_{1})\cdot p(\theta_{1})+f(x|\theta_{2})\cdot p(\theta_{2})+f(x|\theta_{3})\cdot p(\theta_{3}) 
$$


```{r}
pData = sum( pDataGivenTheta * pTheta )
pData
```


Step 3. Let's calculate the posterior for each $\theta$

$$
f(\theta|x)  =\frac{f(x|\theta)\cdot f(\theta)}{f(x)} 
$$


```{r}
pThetaGivenData = pDataGivenTheta * pTheta / pData   
pThetaGivenData
```



Step 4. Lets plot the prior, likelihood, and posterior

```{r}
# Plot the results.
layout( matrix( c( 1,2,3 ) ,nrow=3 ,ncol=1 ,byrow=FALSE ) ) # 3x1 panels
par(mar=c(3,3,1,0))         # number of margin lines: bottom,left,top,right
par(mgp=c(2,1,0))           # which margin lines to use for labels
par(mai=c(0.5,0.5,0.3,0.1)) # margin size in inches: bottom,left,top,right

# Plot the prior:
plot( Theta , pTheta , type="h" , lwd=3 , main="Prior" ,
      xlim=c(0,1) , xlab=bquote(theta) ,
      ylim=c(0,1.1*max(pThetaGivenData)) , ylab=bquote(p(theta)) ,
      cex.axis=1.2 , cex.lab=1.5 , cex.main=1.5 , col="skyblue" )

# Plot the likelihood:
plot( Theta , pDataGivenTheta , type="h" , lwd=3 , main="Likelihood" ,
      xlim=c(0,1) , xlab=bquote(theta) ,
      ylim=c(0,1.1*max(pDataGivenTheta)) , ylab=bquote(paste("p(D|",theta,")")),
      cex.axis=1.2 , cex.lab=1.5 , cex.main=1.5 , col="skyblue" )
text( .55 , .85*max(pDataGivenTheta) , cex=2.0 ,
      bquote( "D=" * .(nHeads) * "H," * .(nTails) * "T" ) , adj=c(0,.5) )

# Plot the posterior:
plot( Theta , pThetaGivenData , type="h" , lwd=3 , main="Posterior" ,
      xlim=c(0,1) , xlab=bquote(theta) ,
      ylim=c(0,1.1*max(pThetaGivenData)) , ylab=bquote(paste("p(",theta,"|D)")),
      cex.axis=1.2 , cex.lab=1.5 , cex.main=1.5 , col="skyblue" )
text( .55 , .85*max(pThetaGivenData) , cex=2.0 ,
      bquote( "p(D)=" * .(signif(pData,3)) ) , adj=c(0,.5) )
```
