---
title: 'Conditional Independence: Berkeley Admission Data'
author: "Le Wang"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(eikosograms)
library(gridExtra)
```


## Visualization of Unconditional and Conditial Results

The visualization example is drawn from [R. W. Oldford's example of eikosograms](https://cran.r-project.org/web/packages/eikosograms/vignettes/DataAnalysis.html) to understand probability and relations between categorical variates. 

Two things stand out:

1. Admission rates differ across gender
2. However, admission rates also differ across departments



```{r}
e1 <- eikos(Admit ~ Gender, data = UCBAdmissions, 
            yaxs = FALSE, xaxs = FALSE, 
            draw = FALSE)
e2 <- eikos(Admit ~ Dept, data = UCBAdmissions,  
            yaxs = FALSE, xaxs = FALSE,
            draw = FALSE)
# Using the gridExtra package, draw these in a single plot
grid.arrange(e1, e2, nrow = 1)
```

What if there is also a gender difference in the type of departments that prospective students apply to?

```{r}
eikos(Gender ~ Dept, data = UCBAdmissions, yprobs = seq(0,1,0.25), xaxs = FALSE)
```



Lets consider all three variables together

```{r}
eikos(Admit ~  Gender + Dept, data = UCBAdmissions,  
      yaxs = FALSE, xaxs = FALSE, 
      lock_aspect = FALSE, 
      xlab_rot = 90, xvals_size = 8,
      ispace = list(bottom = 15))
```


In all departments but C and E (which are near equal), a greater proportion of female applicants were accepted than males.

## R Markdown

Lets obtain a test of **unconditional independence**. First, lets obtain **joint distribution** (frequency counts)

```{r}
berkeley <- as.data.frame(UCBAdmissions)
freq <- berkeley %>% 
  group_by(Admit, Gender) %>% 
  summarise(sum = sum(Freq))

freq


mytable <- xtabs(sum ~ Admit + Gender, data = freq)
prop.table(mytable)

```

Second, we will perform the $\chi^2$ test of independence described before. 

```{r}
chisq.test(mytable)
```

Next, lets perform a test of **conditional independence**. Since we have data in the form of joint distribution already, where the extra dimension `department` is introduced, we are ready to directly perform the test on the obtained joint distribution. 

```{r}
mantelhaen.test(UCBAdmissions)

```

