---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

We will combine information from Census Name data `cnames` and Florida Census data `FLcensus` to predict ethnic outcomes for the Florida voters data, in which we already have information on individual self-reported (true) outcome to evaluate our accuracy. 

## Step 1. First look at the Florida data


We will read the data

```{r}
cnames<-read.csv("data/cnames.csv")
FLvoters<-read.csv("data/FLvoters.csv")
FLcensus<-read.csv("data/FLCensusVTD.csv")
dim(FLvoters)

```


Lets examine the structure of the data (Note that this is individual data that contain the joint distribution and of course we can recover everything that we need.)

```{r}
head(FLvoters)
```


```{r}
head(cnames)
```


```{r}
head(FLcensus)
```

**Question:** What information do we have in each dataset?

$$
\begin{aligned}
& \Pr[\text{Race} \mid \text{Surname}] \\
& \Pr[\text{Surname} \mid \text{Race}] \\
& \Pr[\text{Race} \mid \text{Surname, Residence}] \\
& \Pr[\text{Race} \mid \text{Residence}] \\
\end{aligned}
$$

## Step 2. Clean up the data a bit

### Preliminary: Matching Data

We will drop those data points that are NOT matched to the census names. 

**Approach #1:** Let's learn how to select data using `match()`

```{r}
x<-c("blue","white","red")
y<-c("black","blue")

# Give me the corresponding position in y
match(x,y)

# Give me the corresponding position in x
match(y,x)
```

If I would like to select only x values matched to y, then I will use the following code

```{r}
!is.na(match(x,y))

```


```{r}
x[!is.na(match(x,y))]
```

**Approach #2: Tidy Approach**

```{r}
library(dplyr)
x<-data.frame(x=c("blue","white","red"))
y<-data.frame(y=c("black","blue"))

x %>% 
  filter(x %in% y$y)
```

### Clean Our Data

Keep only those surnames that can be matched to the Census Data

**Approach #1**

```{r}
FLvoters <- FLvoters[!is.na(match(FLvoters$surname,cnames$surname)),]
```

**Approach #2**

```{r}
FLvoters <- FLvoters %>% 
  filter(surname %in% cnames$surname)
```


Let's drop missing values

```{r}
FLvoters<-na.omit(FLvoters)
dim(FLvoters)
```


Lets combine the percentage of other races

```{r}
cnames$pctothers<-100-(cnames$pctapi+cnames$pctblack+cnames$pctwhite+cnames$pcthispanic)
head(cnames)
```

## Step 3: Prediction

$$
\Pr[\text{race} \mid \text{surname}]
$$

1. Split the data by races. For example, whites and blacks. 

```{r}
whites <- subset(FLvoters,subset = (race == "white")) 
head(whites)
```


2. Match the Florida data and the Census data using surname.

```{r}
w.indx <- match(whites$surname, cnames$surname)
head(whites)
head(w.indx)
```


For example, the first person is PIEDRA, and this surname appears as the $8610^{th}$ person in the Census data.

```{r}
head(cnames)
```


```{r}
head(whites)

```


```{r}
cnames[8610,]
```



3. For the sample of whites (census names corresponding to whites), the maximum of conditional probabilities should be the conditional probabilitiy of being white given the surname.

$$
\max\{\text{pctwhite,pctblack,pctapi,pcthispanic}\}=\text{pctwhite}
$$

Example

```{r}
cnames[8610,]
```

```{r}
max(cnames[8610, c("pctwhite","pctblack","pctapi","pctaian","pct2prace","pcthispanic")])
```

Is it equal to the percentage white? Or, is the percentage of whites the largest number? 

```{r}
max(cnames[8610, c("pctwhite","pctblack","pctapi","pctaian","pct2prace","pcthispanic")]) == cnames$pctwhite[8610]
```

Then we know we did not predict correctly. Lets do it for every osbervation then. 

```{r}
# Pick our relevant variables, which should sum to 100 
vars<-c("pctwhite","pctblack","pctapi","pctaian","pct2prace","pcthispanic")

cnames[w.indx, vars]
```


```{r}


#Calculate the maximum and compare it to the cond prob of being
# whites

#apply(cnames[w.indx, vars], 1, max)

head(apply(cnames[w.indx, vars], 1, max) )

comparison<-apply(cnames[w.indx, vars], 1, max) == cnames$pctwhite[w.indx]
head(comparison)

```



Lets look at how we should use `apply()` in R. see mv05_cond_indep03_apply.Rmd




4. The success rates are defined as the instances when these two are indeed the same. 

```{r}
comparison
```


```{r}
mean(comparison)
```


5. We repeat the process (3) and (4) for other races. 


```{r}
# We can repeat the process for other races
# Blacks 

blacks<- subset(FLvoters, subset = (race == "black")) 
b.indx<-match(blacks$surname,cnames$surname) 
mean(apply(cnames[b.indx, vars], 1, max) == cnames$pctblack[b.indx])

```


```{r}
# Asians 
asians <- subset(FLvoters, subset = (race == "asian")) 
a.indx<- match(asians$surname, cnames$surname) 
mean(apply(cnames[a.indx, vars], 1, max) == cnames$pctapi[a.indx])


```


```{r}
# Hispanics 
hispanics <- subset(FLvoters, subset = (race == "hispanic")) 
h.indx <- match(hispanics$surname, cnames$surname) 
mean(apply(cnames[h.indx, vars], 1, max) == cnames$pcthispanic[h.indx])
```



## Step 4: Prediction taking into account more information

Can we improve the precision by incorporating more information, for example, on residence?

$$
 \Pr[\text{race}|\text{surname,residence}]
$$




$$
\Pr[\text{race}|\text{surname,residence}]=\frac{\Pr[\text{surname}|\text{race,residence}]\Pr[\text{race}|\text{residence}]}{\Pr[\text{surname}|\text{residence}]}
$$


What do We have:

$$
\Pr[\text{race}|\text{surname}] \quad \text{(Census data)}
$$

```{r}
head(cnames)
```


$$
\Pr[\text{race}|\text{residence}] \text{ and } \Pr[\text{residence}] \quad \text{(Florida Census data)}
$$
```{r}
head(FLcensus)
```

We do not have in these two datasets (note that `FLvoters` is only our validation dataset):


$$
\Pr[\text{surname}\mid \text{race},\text{residence}] 
$$

$$
\Pr[\text{surname}\mid \text{residence}] 
$$

Note that with the application of the Law of Total Probability

we can obtain the second one 

$$
\begin{aligned}
& \Pr[\text{surname}|\text{residence}]	\\
= & \sum\Pr[\text{surname}|\text{race,residence}]\cdot\Pr[\text{race}|\text{residence}] \\
\end{aligned}
$$
Now the key is to obtain 
$$
\Pr[\text{surname}\mid \text{race,residence}] 
$$

We would invoke our conditional independence assumption

$$
\Pr[\text{surname} \mid \text{race, residence}]=\Pr[\text{surname}\mid \text{race}]
$$

Does it make sense? Or, when will it be violated? Conditional independence implies that once we know a voter's race, her residence location does not give us any additional information about her surname. 

There is NO strong geographical concentration of certain surnames in Florida within a racial categroy. 

$$
\begin{aligned}
& \Pr[\text{surname}|\text{race,}\text{residence}]	\\
\\
\\
= & \Pr[\text{surname}|\text{race}] \\
\\
\\
= & \frac{\Pr[\text{race}|\text{surname}]\Pr[\text{surname}]}{\Pr[\text{race}]}
\end{aligned}
$$
We can obtain these terms:

1. Census data: 
$$\Pr[\text{race}|\text{surname}] \text{ and } \Pr[\text{surname}]$$ 

```{r}
head(cnames)
```


2. Florida Census data:
$$
\Pr[\text{race}]=\sum_{\text{residence}}\Pr[\text{race}|\text{residence}]\Pr[\text{residence}]
$$
```{r}
head(FLcensus)
```


```{r}
race.prop <- apply(FLcensus[,c("white", "black", "api", "hispanic", "others")], 
					2, 
					weighted.mean, 
					weights = FLCensus$total.pop)
race.prop
```


So, if we are interested in 
$$
\Pr[\text{surname} \mid \text{race=white}]=\frac{\Pr[\text{white} \mid \text{surname}]\Pr[\text{surname}]}{\Pr[\text{race=white}]}
$$

```{r}
head(cnames)
total.count<- sum(cnames$count)

cnames$name.white <- (cnames$pctwhite/100)*(cnames$count/total.count)/race.prop["white"]
```


```{r}
head(cnames)
```


```{r}
cnames$name.black <- (cnames$pctblack/100)*(cnames$count/total.count)/race.prop["black"]

cnames$name.hispanic <- (cnames$pcthispanic/100)*(cnames$count/total.count)/race.prop["hispanic"]

cnames$name.api <- (cnames$pctapi/100)*(cnames$count/total.count)/race.prop["api"]

cnames$name.others <- (cnames$pctothers/100)*(cnames$count/total.count)/race.prop["others"]

```

Lets look at what we have now

```{r}
head(cnames)
```

Now let's get the following information

$$
\begin{aligned}
& \Pr[\text{surname}|\text{residence}]	\\
= & \sum\Pr[\text{surname}|\text{race,residence}]\cdot\Pr[\text{race}|\text{residence}] \\
= & \sum\Pr[\text{surname}|\text{race}]\cdot\Pr[\text{race}|\text{residence}] 
\end{aligned} 
$$

Lets merge all the information with our validation dataset

```{r}
FLvoters<- merge(x=FLvoters, y=FLcensus, by=c("county","VTD"), all= FALSE)

head(FLvoters)
```


```{r}
indx <- match(FLvoters$surname, cnames$surname) 

FLvoters$name.residence <-cnames$name.white[indx]*FLvoters$white+cnames$name.black[indx]*FLvoters$black+cnames$name.hispanic[indx]*FLvoters$hispanic+cnames$name.api[indx]*FLvoters$api+cnames$name.others[indx]*FLvoters$others 
head(FLvoters)
```


Finally,

$$
\begin{aligned}
 & \Pr[\text{race}|\text{surname,residence}] \\	
\\
 = & \frac{\Pr[\text{surname}|\text{race,residence}]\Pr[\text{race}|\text{residence}]}{\Pr[\text{surname}|\text{residence}]} \\
 \\
&	= \frac{\Pr[\text{surname}|\text{race}]\Pr[\text{race}|\text{residence}]}{\Pr[\text{surname}|\text{residence}]}

\end{aligned}
$$

For example, 
$$
\begin{aligned}
& \Pr[\text{white}|\text{surname,residence}] \\
\\
= & \frac{\Pr[\text{surname}|\text{white}]\Pr[\text{white}|\text{residence}]}{\Pr[\text{surname}|\text{residence}]}
\end{aligned}
$$

```{r}
FLvoters$predict.white <- cnames$name.white[indx]*FLvoters$white/FLvoters$name.residence
FLvoters$predict.black <- cnames$name.black[indx]*FLvoters$black/FLvoters$name.residence
FLvoters$predict.hispanic <- cnames$name.hispanic[indx]*FLvoters$hispanic/FLvoters$name.residence
FLvoters$predict.api <- cnames$name.api[indx]*FLvoters$api/FLvoters$name.residence
FLvoters$predict.others <- cnames$name.others[indx]*FLvoters$others/FLvoters$name.residence
```

```{r}
## relevant variables 
vars1 <- c("predict.white", "predict.black", "predict.hispanic", "predict.api",            "predict.others") 

## whites 
whites <- subset(FLvoters, subset = (race == "white")) 
mean(apply(whites[, vars1], 1, max) == whites$predict.white) 

## blacks 
blacks <- subset(FLvoters, subset = (race == "black")) 
mean(apply(blacks[, vars1], 1, max) == blacks$predict.black)
```


