---
title: "Linear Regression and Prediction and Partial Effects (Binary Predictor)"
output: html_notebook
---


This program illustrates how to estimate linear regression and calculate the CEF for two different purposes


```{r}
wagedata <- read.csv("data/wage2.csv") 

# CEF for each category
  aggregate(wage,by=list(married),FUN=mean) 
  results <-aggregate(wage,by=list(married),FUN=mean)[2] 
```

We also calculated the partial effects

```{r}
  results[2,1]-results[1,1]
```




# Linear Regression without intercept

$$
\text{wage}=\beta_{1}\cdot\text{non-married}+\beta_{2}\cdot\text{married}+\epsilon
$$

How to implement this in R?

```{r}
lm(wage ~ factor(married)+0, data=wagedata)
```

Or, you can use R's built-in command to summarize the results 

```{r}
results <- lm(wage ~ factor(married)+0, data=wagedata)
# Linear Regression without intercept
summary(results)
```

This model implies that the prediction will be

$$
\begin{aligned}
\text{wage}&=\beta_{1}\cdot\text{non-married}+\beta_{2}\cdot\text{married}+\epsilon\\
 &=798.4400\cdot \text{non-married}+977.0479\cdot\text{married}
 \end{aligned}
$$

We can use R to make nice tables

```{r}
# Use R to make nice tables
library(knitr)
kable(summary(results)$coef)
```

If you are not sure what the coefficients are..let's repeat what we have done earlier. 
```{r}
# Obtain Conditional Means
aggregate(wage,by=list(married),FUN=mean) 
```


What is your prediction if someone is not married? What if someone is married?

```{r}
# As we did with Naive Bayes Models
new <- data.frame(married=c(0,1)) 
predict(results,new)

aggregate(wage,by=list(married),FUN=mean)
```

What is the partial impact of marriage on wages? The same!

# Linear Regression with intercept

Let's make another model using the exact same information:

$$
\text{wage}=\beta_{0}+\beta_{1}\cdot\text{married}+\epsilon
$$




```{r}
model <- lm(wage ~ married) 
summary(model)
```

These estimates imply the following model

$$
\text{wage}=798.44+ 178.61 \cdot\text{married}
$$

If you recall, the intercept is exactly the mean wage for non-married individuals, and the slope coefficient on married is exactly the partial effect (i.e., the difference in the mean wages between married and non-married individuals)!

What are the predictions from this model?

```{r}
# Generate Predictions
new <- data.frame(married=c(0,1)) 
predict(model,new)
```


```{r}
# Obtain Conditional Means
aggregate(wage,by=list(married),FUN=mean)
```





