---
title: "K-Nearest Neighbor Estimation"
output: html_notebook
---

My code works for a single variable with Eucledian distance as a distance measure. But in practice, we will have more than one predictor variables and use other distance measures. And more important, my codes does not take into account ties. So, we can use some built-in commands

1. `knn.reg()` in the package `FNN`
2. `knn()` in the package `class`

## K-nearest Neighbor Regression 

```{r}
# Install necessary package if you have not done so
# install.packages("FNN")
library(FNN)
```


```{r}
wagedata <- read.csv("data/wage2.csv") 
attach(wagedata)
```


To illustrate the limitation of my code and how to use the command `knn.reg`

Note that you need to load my `KKN` function before runing this block of code.

```{r}
mydata <- data.frame(IQ = c(50, 80, 100))
knn.reg(train = wagedata$IQ, test = mydata, y=wagedata$wage, k = 3)

mydata <- c(50, 80, 100)
sapply(mydata,KKN,K=3)
```

Why do the results not match, except for the first data point?

```{r}
IQ[order(abs(IQ-80))]
```

What will happen if we just pick the first three observations?

```{r}
IQ[order(abs(IQ-80))][1:3]
```

This is a limitation of my code and not a desirable property. We can use `knn.reg()` in practice as follows


```{r eval=FALSE}
knn.reg(train = ?, test = ?, y = ?, k = ?)
```

Note that 

1. `train`: the predictors of the training data
2. `test`: the predictor values,  $x$, at which we would like to make predictions
3. `y`: the response for the training data
4. `k`: the number of neighbors

Lets make a prediction for the following individuals

```{r}
mydata <- data.frame(IQ = c(50, 80, 100), exper = c(7,8,11))
mydata
```

Lets estimate the model and make predictions
```{r}
test.data <- data.frame(IQ = IQ, exper = exper)
knn.reg(train = test.data, test = mydata, y=wage, k = 3)
```

## K-nearest Neighbor Classification

Lets assume that our goal is to predict (or classify) whether or not someone is married. We just find the most likely outcome (i.e., the distribution of the outcome) among the k-nearest neighbors and use it as our prediction. 

```{r}
# Install necessary package if you have not done so
# install.packages("class")
library(class)
```

We can use `knn()` in practice as follows


```{r eval=FALSE}
knn(train = ?, test = ?, cl = ?, k = ?)
```

Note that 

1. `train`: the predictors of the training data
2. `test`: the predictor values,  $x$, at which we would like to make predictions
3. `cl`: the true class labels for the train set
4. `k`: the number of neighbors

Lets make a prediction for the following individuals

```{r}
mydata <- data.frame(IQ = c(50, 80, 100), exper = c(7,8,11))
mydata
```

Lets estimate the model and make predictions
```{r}
test.data <- data.frame(IQ = IQ, exper = exper)
knn(train = test.data, test = mydata, cl=married, k = 3, prob = TRUE)
```

