---
title: "Nonparametric Kernel Estimation"
output: html_notebook
---

```{r}
wagedata <- read.csv("data/wage2.csv") 
```

```{r}
# Install the package if you have not done so
# install.package("np")
library(np)
```


## Step 1.
Lets estimate the model first

```{r}
model.np <- npreg(wage ~ IQ + exper,
                  data = wagedata)
```

## Step 2. 
Lets make the dataset for which we would like to make the predictions
```{r}
mydata <- data.frame(IQ = c(50, 80, 100), exper = c(7,8,11))
mydata
```

## Step 3.
Using the estimated model, we will make the predictions

```{r}
predict(model.np, newdata = mydata)
```

# In the presence of discrete variables (ordered or unordered)

## Step 1.

Note that when you have discrete variables, some are ordered and others unordered, it is important to take into account the nature of such variables. 

```{r}
model.np <- npreg(wage ~ ordered(IQ) 
                  + factor(married)
                  + exper,
                  data = wagedata)
```

## Step 2. 
Lets make the dataset for someone with IQ of 50, experience of 7 years, and currently being married. 

```{r}
mydata <- data.frame(IQ = c(50), exper = c(7), married = c(1))
mydata
```

## Step 3.
Using the estimated model, we will make the predictions

```{r}
predict(model.np, newdata = mydata)
```
