---
title: "Introduction to Writing R Functions"
author: "Le Wang"
date: "4/15/2017"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Motivation

As in most programming languages, the heart of `R` programming consists of writing *functions*.  

One purpose of writing an function is to avoid or simplify repetitive tasks. It is more elegant and cleaner.  

## Differences between Stata and R

In Stata, we *usually* don’t write functions; we execute built-in commands like `browse` or `gen` or `logit`. You can write an `.ado` file to make a new command. 

In R, there are built-in functions like `summary()` or `glm()` or `median()`, but you can also write your own functions. 

## Motivation: Example

Suppose that we would like to calculate the mean of a variable or even more variables.

```{r wasup, echo=FALSE}
set.seed(123456)
x1<-rnorm(100)
x2<-rnorm(100)
x3<-rnorm(100)
```

```{r wasup2, size='footnotesize'}
N   <- length(x1)
sum <- sum(x1)
mean <- sum/N

N   <- length(x2)
sum <- sum(x2)
mean <- sum/N

N   <- length(x3)
sum <- sum(x3)
mean <- sum/N
```

## Motivation

Anticipating that we may have to repeat this process many times in the future, we can write a function to execute the whole thing for us. 

A function is a group of **instructions** that **take inputs**, uses them to **compute other values**, and **returns a result**. It usually takes the following form

```{r, eval=FALSE, size='footnotesize'}
mymean <- function(x){
  
  instructions to compute certain values
  
  return(some values)
  
}
```

Then, you call the function for your own case by typing 

```{r, eval=FALSE}
mymean(x1)
```

## Distinct Elements in an R function

```{r, eval=FALSE, size='footnotesize'}
mymean <- function(x){
  
  instructions to compute certain values
  
  return(some values)
  
}
```

1. Function name
2. Required Input (**formal arguments**). They are simply *placeholders* for **actual arguments**.
3. Instructions or Body of Code

## First Step: What is common and What is not

Perform the task for sepcific cases several times to detect 

1. What changes: usually would be the **input** that comes from an user. 
2. What does not: usually would be the body inside the function.  

## First Step:

```{r, size='footnotesize'}
N   <- length(x1)
sum <- sum(x1)
mean <- sum/N

N   <- length(x2)
sum <- sum(x2)
mean <- sum/N

N   <- length(x3)
sum <- sum(x3)
mean <- sum/N
```

1. What changes? **Variable name**
2. What does not change? The procedure to calculate the mean

## Second Step: Program Structure

```{r, eval=FALSE}
mymean <- function(){

  # Calculate the mean
  
  
  # Return the mean
  return(mean)
}
```


## Third Step: Filling in the main code

```{r, eval=FALSE}
mymean <- function(){

  # Calculate the mean
  N   <- length(_____)
  sum <- sum(____)
  mean <- sum/N  
  
  # Return the mean
  return(mean)
}
```

## Fourth Step: Giving a placeholder

```{r, eval=FALSE}
mymean <- function(var){

  # Calculate the mean
  N   <- length(var)
  sum <- sum(var)
  mean <- sum/N  
  
  # Return the mean
  return(mean)
}
```

## Fourth Step: Giving a placeholder

```{r, eval=FALSE}
mymean <- function(x){

  # Calculate the mean
  N   <- length(x)
  sum <- sum(x)
  mean <- sum/N  
  
  # Return the mean
  return(mean)
}
```

## Final Step: Use it!

```{r, size='footnotesize'}
mymean <- function(x){

  # Calculate the mean
  N   <- length(x)
  sum <- sum(x)
  mean <- sum/N  
  
  # Return the mean
  return(mean)
}

mymean(x1)
mean(x1)
```

## Final Step: Use it!
```{r,size='footnotesize'}
mymean(x2)
mean(x2)

mymean(x3)
mean(x3)
```

## Example 2: KNN Estimator

**Goal**: 

\bigskip

We would like to use the average wages of the K-nearest neighors for my prediction. 

\bigskip
**Translations:**

1. Find out the K-nearest neighbors
2. Calculate the average wages. 

## Example 2: KNN Estimator

**Question**: How to find out the K-nearest neighbors

1. Distance measures between all the observations and your value $x$
2. Find out where the closest $K$ values are (index/position)

```{r,echo=FALSE, message=FALSE}
library(readr)
wage2 <- read_csv("data/wage2.csv")
attach(wage2)
````

```{r}
# Distance
d <- abs(IQ-60)
# List the data from the smallest
order(d)[1:3]
neighbor<-order(d)[1:3]
mean<-mean(wage[neighbor])
mean
```

## Example 2: KNN Estimator
Let's see more special cases and find out what can change

```{r}
# Distance
d <- abs(IQ-80)
# List the data from the smallest
order(d)[1:3]
neighbor<-order(d)[1:3]
mean<-mean(wage[neighbor])
mean
```

## Example 2: KNN Estimator
Let's see more special cases and find out what can change

```{r}
# Distance
d <- abs(IQ-100)
# List the data from the smallest
order(d)[1:3]
neighbor<-order(d)[1:3]
mean<-mean(wage[neighbor])
mean
```

## Example 2: KNN Estimator
Let's see more special cases and find out what can change

```{r}
# Distance
d <- abs(IQ-100)
# List the data from the smallest
order(d)[1:5]
neighbor<-order(d)[1:5]
mean<-mean(wage[neighbor])
mean
```

## Example 2: KNN Estimator

```{r}
# Distance
d <- abs(IQ-80)
# List the data from the smallest
order(d)[1:3]
neighbor<-order(d)[1:3]
mean<-mean(wage[neighbor])
mean
```

What can change? 

1. The value of the predictor that you calculate the conditional mean for 
2. The definition of *closedness*: $K$. 

## Example 2: KNN Estimator

```{r,eval=FALSE}
# Distance
d <- abs(IQ-_______)
# List the data from the smallest
order(d)[1:_______]
neighbor<-order(d)[1:_______]
mean<-mean(wage[neighbor])
mean
```

What can change? 

1. The value of the predictor that you calculate the conditional mean for 
2. The definition of *closedness*: $K$. 

## Step 2: Program Structure

```{r, eval=FALSE}
myKNN <- function(){

  # Calculate the mean
  
  
  # Return the mean
  return(mean)
}
```

## Step 3: Filling in the main code

```{r, eval=FALSE}
myKNN <- function(){

  # Distance
  d <- abs(IQ-_______)
  # List the data from the smallest
  order(d)[1:_______]
  neighbor<-order(d)[1:_______]
  mean<-mean(wage[neighbor])  
  
  # Return the mean
  return(mean)
}
```

## Step 4: Giving it Two Placeholders!

```{r, eval=FALSE}
myKNN <- function(x,K){

  # Distance
  d <- abs(IQ-x)
  # List the data from the smallest
  order(d)[1:K]
  neighbor<-order(d)[1:K]
  mean<-mean(wage[neighbor])  
  
  # Return the mean
  return(mean)
}
```

## Final Step: Use it!

```{r}
myKNN <- function(x,K){

  # Distance
  d <- abs(IQ-x)
  # List the data from the smallest
  order(d)[1:K]
  neighbor<-order(d)[1:K]
  mean<-mean(wage[neighbor])  
  
  # Return the mean
  return(mean)
}
myKNN(60,3)
```

## Applications of What we use

e.g., 

`order()` can be used to pick out your recommendation for Airnb clients.

`order()` can be used to pick out the furthest neighbors. (How to change the option so that the results are in the decreasing order?)

## Quiz: Question 1

Writen an `R` function to calculate the square of any number and return it. Call this function `square.it()`

```{r}
1^2
2^2
3^2
4^2
```

```{r}

```


## Quiz: Question 2 

Write an `R` function to calculate sample variance. Call this function `myvar()`
$$
\frac{1}{N-1} \sum (x_i-\bar{x})^2
$$
```{r}
set.seed(123456)
x<-rnorm(1000)
var(x)
```

```{r}
myvar <- function(){
  
  # calculate sample mean
  
  # construct a new variable that is the square 
  # of the difference between a variable and its sample mean
  
  # calculate sum
  
  # Calculate the lenght of the data, N
  
  # Divide sum by N-1
  
  var <-
  
  
  return(var)
}

myvar(x1)

```

## Building More Complex R functions

We can build on what we have to create new functions. 

\bigskip 

A function that calls another function that calls another function and so on. 
\bigskip

**Task:** Suppose that we would like to build a function for **sample conditional variance**

## Goals

Remember that **conditional** is nothing but a sub-sample. We can select the sub-sample and then calculate the sample variance using our code for sample variance. 

```{r}
aggregate(wage,by=list(black),FUN=var)

# Black
aggregate(wage,by=list(black),FUN=var)[1,2]

# Non-Black
aggregate(wage,by=list(black),FUN=var)[2,2]
```

## Conditional Statement in R

```{r, eval=FALSE}
if (condition 1) {
    statement 1
} else if (condition 2) {
    statement 2
} else {
    statement 3
}
```

## Conditional Statement in R

```{r, eval=FALSE}
if (an indivdidual is black) {
    variance for black
} else {
    variance for non-black
}
```

## Conditional Statement in R

```{r, eval=FALSE}
aggregate(wage,by=list(black),FUN=var)

if (_______ == 1) {
  aggregate(wage,by=list(black),FUN=var)[2,2]
} else {
  aggregate(wage,by=list(black),FUN=var)[1,2]
}
```


## Conditional Variance

```{r}
mycvar <- function(x){
  
  aggregate(wage,by=list(black),FUN=var)
  
  if (x == 1) {
    cvar<-aggregate(wage,by=list(black),FUN=var)[2,2]
  } else {
    cvar<-aggregate(wage,by=list(black),FUN=var)[1,2]
  }  
  
  return(cvar)
}

mycvar(1)

```

